---
title: "20250603gutenberg"
output: html_document
---

```{r libraries}
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(lubridate)
library(ggforce)
library(lme4)
```


```{r getdata}
gutenberg_authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-03/gutenberg_authors.csv')
gutenberg_languages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-03/gutenberg_languages.csv')
gutenberg_metadata <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-03/gutenberg_metadata.csv')
gutenberg_subjects <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-03/gutenberg_subjects.csv')

```

```{r glimpse_all}
glimpse(gutenberg_authors)
glimpse(gutenberg_languages)
glimpse(gutenberg_metadata)
glimpse(gutenberg_subjects)
```
#How many different languages are available in the Project Gutenberg collection? How many books are available in each language?

```{r frequency}
gutenberg_languages %>%
  group_by(language) %>%
  summarise(count = n())
```

##There are 70 different languages in the Project Gutenberg Collection. 
The above tibble shows how many are available in each language. Most frequent being English with 60,693 books.

#Do any authors appear under more than one gutenberg_author_id?

```{r count_authors}
gutenberg_authors %>%
  group_by(author) %>%
  summarise(count = n()) %>%
  filter(count > 1)
```

##A total of 119 authors have more than one author id.

# How might the {gutenbergr} package authors further refine the data for greater ease-of-use?


